### 2018-01-11 submission/manuscript.Rmd
update RF section except one paragraph

***Whole Community Models Add Important Context:***
Since CRC-associated genera increased RR for carcinoma over diversity metrics we assessed whether the bacterial community was better at classifying disease versus only CRC-associated genera and tested this using two approaches. The first approach used genus level data and tested whether there were any differences in AUC between all genera and CRC-associated genera only during training on one study or testing on all other studies. The second approach used OTU level data and tested whether there was a generalized decrease in the 10-fold cross validation (CV) model AUC across studies when comparing all OTUs versus only OTUs that taxonomically classified to CRC-associated genera.

@@

In stool the most common genera in the genus based models belonged to genera such as *Ruminococcus*, *Bacteroides*, and *Roseburia* [Figure 6A & B]. Although all four CRC-associated genera were present in carcinoma, none were present in the majority of studies and *Fusobacterium* was the only genus present in adenoma [Figure 6A & B]. For the OTU based models, *Ruminococcaceae* was present in the top 10 consistently for both adenoma and carcinoma models while *Roseburia* was only present in many adenoma models and *Bacteroides* was present in the overwhelming majority of the carcinoma models [Figure 6C & 6D].

Unlike stool, tissue based Random Forest models showed no consistent representation of *Ruminococcaceae*, *Ruminococcus*, *Bacteroides*, and *Roseburia* in the top 10 across study [Figure S7]. The vast majority of the top 10 genera and OTUs occured in a study specific manner for these tissue based Random Forest models. For both the genera and OTU Random Forest models of adenoma and carcinoma, there appears to be very little overlap in the top 10 most important variables between stool and tissue [Figure 6 & S7]. This discordance between stool and tissue also applies to the CRC-associated genera with one noticeable skew being that *Fusobacterium* and *Fusobacteriaceae* occur more often in the top 10 of Random Forest models using matched versus unmatched tissue samples [Figure S7B-C & S7E-F].

>>>>

***Whole Community Models Add Important Context:***
Since specific genera increased RR for carcinoma over diversity metrics we assessed whether the bacterial community was better at classifying disease versus only a select group of genera based on RR and P-value significance and tested this using two approaches. The first approach used genus level data and tested whether there were any differences in AUC between all genera and selected genera during training on one study or testing on all other studies. The second approach used OTU level data and tested whether there was a generalized decrease in the 10-fold cross validation (CV) model AUC across studies when comparing all OTUs versus only OTUs that taxonomically classified to selected genera.

@@

In stool the most common genera in the full genera-based models belonged to genera such as *Ruminococcus*, *Bacteroides*, and *Roseburia* [Figure 6A & B]. Although all four CRC-associated genera were present in carcinoma, none were present in the majority of studies and *Fusobacterium* was the only genus present in adenoma [Figure 6A & B]. For the full OTU-based models, *Ruminococcaceae* was present in the top 10 consistently for both adenoma and carcinoma models while *Roseburia* was only present in many adenoma models and *Bacteroides* was present in the overwhelming majority of the carcinoma models [Figure 6C & 6D].

Unlike stool, the tissue full genera-based Random Forest models showed no consistent representation of *Ruminococcaceae*, *Ruminococcus*, *Bacteroides*, and *Roseburia* in the top 10 across study [Figure S7]. The vast majority of the top 10 genera and OTUs occured in a study specific manner for these tissue based Random Forest models. For both the full genera and OTU Random Forest models of adenoma and carcinoma, there appears to be very little overlap in the top 10 most important variables between stool and tissue [Figure 6 & S7]. This discordance between stool and tissue also applies to the mouth-associated genera with one noticeable skew being that *Fusobacterium* and *Fusobacteriaceae* occur more often in the top 10 of Random Forest models using matched versus unmatched tissue samples [Figure S7B-C & S7E-F].

### 2018-01-11 submission/manuscript.tex
update RF section except one paragraph

\textbf{\emph{Whole Community Models Add Important Context:}} Since
CRC-associated genera increased RR for carcinoma over diversity metrics
we assessed whether the bacterial community was better at classifying
disease versus only CRC-associated genera and tested this using two
approaches. The first approach used genus level data and tested whether
there were any differences in AUC between all genera and CRC-associated
genera only during training on one study or testing on all other
studies. The second approach used OTU level data and tested whether
there was a generalized decrease in the 10-fold cross validation (CV)
model AUC across studies when comparing all OTUs versus only OTUs that
taxonomically classified to CRC-associated genera.

@@

In stool the most common genera in the genus based models belonged to
genera such as \emph{Ruminococcus}, \emph{Bacteroides}, and
\emph{Roseburia} {[}Figure 6A \& B{]}. Although all four CRC-associated
genera were present in carcinoma, none were present in the majority of
studies and \emph{Fusobacterium} was the only genus present in adenoma
{[}Figure 6A \& B{]}. For the OTU based models, \emph{Ruminococcaceae}
was present in the top 10 consistently for both adenoma and carcinoma
models while \emph{Roseburia} was only present in many adenoma models
and \emph{Bacteroides} was present in the overwhelming majority of the
carcinoma models {[}Figure 6C \& 6D{]}.

Unlike stool, tissue based Random Forest models showed no consistent
representation of \emph{Ruminococcaceae}, \emph{Ruminococcus},
\emph{Bacteroides}, and \emph{Roseburia} in the top 10 across study
{[}Figure S7{]}. The vast majority of the top 10 genera and OTUs occured
in a study specific manner for these tissue based Random Forest models.
For both the genera and OTU Random Forest models of adenoma and
carcinoma, there appears to be very little overlap in the top 10 most
important variables between stool and tissue {[}Figure 6 \& S7{]}. This
discordance between stool and tissue also applies to the CRC-associated
genera with one noticeable skew being that \emph{Fusobacterium} and
\emph{Fusobacteriaceae} occur more often in the top 10 of Random Forest
models using matched versus unmatched tissue samples {[}Figure S7B-C \&
S7E-F{]}.

>>>>

\textbf{\emph{Whole Community Models Add Important Context:}} Since
specific genera increased RR for carcinoma over diversity metrics we
assessed whether the bacterial community was better at classifying
disease versus only a select group of genera based on RR and P-value
significance and tested this using two approaches. The first approach
used genus level data and tested whether there were any differences in
AUC between all genera and selected genera during training on one study
or testing on all other studies. The second approach used OTU level data
and tested whether there was a generalized decrease in the 10-fold cross
validation (CV) model AUC across studies when comparing all OTUs versus
only OTUs that taxonomically classified to selected genera.

@@

In stool the most common genera in the full genera-based models belonged
to genera such as \emph{Ruminococcus}, \emph{Bacteroides}, and
\emph{Roseburia} {[}Figure 6A \& B{]}. Although all four CRC-associated
genera were present in carcinoma, none were present in the majority of
studies and \emph{Fusobacterium} was the only genus present in adenoma
{[}Figure 6A \& B{]}. For the full OTU-based models,
\emph{Ruminococcaceae} was present in the top 10 consistently for both
adenoma and carcinoma models while \emph{Roseburia} was only present in
many adenoma models and \emph{Bacteroides} was present in the
overwhelming majority of the carcinoma models {[}Figure 6C \& 6D{]}.

Unlike stool, the tissue full genera-based Random Forest models showed
no consistent representation of \emph{Ruminococcaceae},
\emph{Ruminococcus}, \emph{Bacteroides}, and \emph{Roseburia} in the top
10 across study {[}Figure S7{]}. The vast majority of the top 10 genera
and OTUs occured in a study specific manner for these tissue based
Random Forest models. For both the full genera and OTU Random Forest
models of adenoma and carcinoma, there appears to be very little overlap
in the top 10 most important variables between stool and tissue
{[}Figure 6 \& S7{]}. This discordance between stool and tissue also
applies to the mouth-associated genera with one noticeable skew being
that \emph{Fusobacterium} and \emph{Fusobacteriaceae} occur more often
in the top 10 of Random Forest models using matched versus unmatched
tissue samples {[}Figure S7B-C \& S7E-F{]}.
